Management induced changes of soil organic carbon on global croplands

### 2022-06-14 ResultNotebooks/results_SOCscenario.Rmd

rev         <- "rev26"
scenarios   <- c("rev26histManagement", "rev26constManagement", "rev26constManagement2", "rev26constManagement_constClimate", "rev26constManagement2_constClimate", "rev26constManure", "rev26constResidues", "rev26constResrate", "rev26constTillage", "rev26constYield", "rev26Initial-1901_constManagement", "rev26Initial-1901_constManagement2", "rev26Initial-1901", "rev26Initial-0_constManagement", "rev26Initial-0_constManagement2", "rev26Initial-0")

# filesNeeded  <- c("SoilCarbon.rds","Landuse.rds")
# setConfig(outputfolder = "/home/kristine/mnt/rd3mod/inputdata/output/")
# 
# untar(paste0(getConfig("outputfolder"),"/", rev, "histManagement_h12_carbonbudget.tgz"),
#       exdir=paste0("Data/",rev,"histManagement"), extras="--keep-newer-files ")

# for(item in scenarios){
#    untar(paste0(getConfig("outputfolder"),"/", paste0(rev,item), "_h12_carbonbudget.tgz"),
#    files=filesNeeded, exdir=paste0("Data/", rev, item), extras="--keep-newer-files ")
# }

@@

for(scen in scenarios){
  SoilCarbon <- read.magpie(paste0("Data/", scen, "/SoilCarbon.rds"))
  Landuse    <- read.magpie(paste0("Data/", scen, "/Landuse.rds"))
  SoilCarbon[,,"naturalstate.crop"] <- SoilCarbon[,,"naturalstate.natveg"]
  SOCStock        <- toolConditionalReplace(dimSums(dimSums(SoilCarbon[,,"state",pmatch=TRUE], dim=3.3)* Landuse, dim=3.2), "is.nan()",0)
  glo_SOCStock    <- mbind(glo_SOCStock, add_dimension(dimSums(SOCStock, dim=1), 
                                                       dim=3.1, nm=scen, add="scenario")) 
}

@@

CarbonInput  <- read.magpie(paste0("Data/rev26histManagement/CarbonInput.rds"))
Landuse      <- read.magpie(paste0("Data/rev26histManagement/Landuse.rds"))
glo_SOCInput <- collapseNames(dimSums(CarbonInput[,,"totalC"] * Landuse, dim=c(1)))/collapseNames(dimSums(Landuse, dim=c(1)))

@@

manage_scen <-  c(histManagement   = "rev26histManagement",
                  constTillage     = "rev26constTillage", 
                  constManure      = "rev26constManure",
                  constResidues    = "rev26constResidues", 
                  constManagement  = "rev26constManagement")#,
                  #constManagementNoCC = "rev26constManagement_constClimate")

@@

limits <- c(-40,-32)
breaks <- c(-32, -34,- 36, -38, -40)

>>>>

rev         <- "rev30"
scenarios   <- c("histManagement", "constClimate-1975", "constManagement-1975", "constManagement2-1975", "constManagement-1975_constClimate-1975", "constManagement2-1975_constClimate-1975", "constManure-1975", "constResidues-1975", "constResrate-1975", "constTillage-mixedtill", "constYield-1975")

 filesNeeded  <- c("SoilCarbon.rds","Landuse.rds")
# setConfig(outputfolder = "/home/kristine/WORK/Paper/historicalsocbudget/ResultNotebooks/Data/")

#  untar(paste0(getConfig("outputfolder"),"/", rev, "histManagement_h12_carbonbudget.tgz"),
#        exdir=paste0("Data/",rev,"histManagement"), extras="--keep-newer-files ")
# 
# for(item in scenarios){
#     untar(paste0(getConfig("outputfolder"),"/", paste0(rev,item), "_h12_carbonbudget.tgz"),
#     exdir=paste0("Data/", rev, item), extras="--keep-newer-files ")
#  }

@@

for(scen in scenarios){
  print(scen)
  SoilCarbon <- read.magpie(paste0("Data/", paste0(rev,scen), "/SoilCarbon.rds"))
  Landuse    <- read.magpie(paste0("Data/", paste0(rev,scen), "/Landuse.rds"))
  SoilCarbon[,,"naturalstate.crop"] <- SoilCarbon[,,"naturalstate.natveg"]
  SOCStock        <- toolConditionalReplace(dimSums(dimSums(SoilCarbon[,,"state",pmatch=TRUE], dim=3.3)* Landuse, dim=3.2), "is.nan()",0)
  glo_SOCStock    <- mbind(glo_SOCStock, add_dimension(dimSums(SOCStock, dim=1), 
                                                       dim=3.1, nm=scen, add="scenario")) 
}

@@

CarbonInput  <- read.magpie(paste0("Data/", rev, "histManagement/CarbonInput.rds"))
Landuse      <- read.magpie(paste0("Data/", rev, "histManagement/Landuse.rds"))
glo_SOCInput <- collapseNames(dimSums(CarbonInput[,,"totalC"] * Landuse, dim=c(1)))/collapseNames(dimSums(Landuse, dim=c(1)))

@@

manage_scen <-  c(histManagement   = "histManagement",
                  constTillage     = "constTillage-mixedtill", 
                  constManure      = "constManure-1975",
                  constResidues    = "constResidues-1975", 
                  constManagement  = "constManagement-1975")#,
                  #constManagementNoCC = "rev26constManagement_constClimate")

@@

limits <- c(-42,-34)
breaks <- c(-34, -36,- 38, -40, -42)

### 2022-06-14 ResultNotebooks/results_SOCtable.Rmd

scenarios <- c("rev26histManagement") #,"rev18histManagement_ResAllom_2021_02_01")
outputdir <- "Output/"

@@

years             <- findset("past")
climateIPCCr      <- readRDS("Data/rev26histManagement/IPCC_reduced.rds")

@@

CShare_ipcczone   <- dimSums(climateIPCCr * SOCDens[,,"crop"][,,"actualstate"]   * Landuse[,years,"crop"], dim=1)/
                     dimSums(climateIPCCr * SOCDens[,,"natveg"][,,"naturalstate"] * Landuse[,years,"crop"], dim=1)

# CShare_ipcczone   <- dimSums(climateIPCCr * toolConditionalReplace(SOCDens[,,"crop"][,,"actualstate"]/SOCDens[,,"natveg"][,,"naturalstate"],
#                                                                  c("is.na()", "is.nan()", "is.infinite()",">3"), c(1,1,1,5)) * Landuse[,years,"crop"], dim=1)/
#                     dimSums(climateIPCCr * Landuse[,years,"crop"], dim=1)

getNames(CShare_ipcczone) <- gsub("_", " ", getNames(CShare_ipcczone))

@@

climateIPCCr       <- readRDS("Data/rev26histManagement/IPCC_reduced.rds")
climateIPCC        <- readRDS("Data/rev26histManagement/IPCC.rds")

@@

MShare_ipcc <- LShare_ipcc <-  HShare_ipcc <- new.magpie("GLO",NULL,outer(c("IPCC2006","IPCC2019"),
                                                                          getNames(climateIPCCr), paste, sep="."))

@@

comparison <- mbind(CShare_ipcc, 
                    round(add_dimension(setYears(collapseNames(CShare_ipcczone[,"y1990","naturalstate"]),NULL),
                                  nm="This Study.hist.1990",add="model.input.year"),2),
                    round(add_dimension(setYears(collapseNames(CShare_ipcczone[,"y2010","naturalstate"]),NULL),
                                  nm="This Study.hist.2010",add="model.input.year"),2))

@@

save_kable(save, file="Output/Images/TableSCF_comparison.png")
table
```

>>>>

scenarios <- c("rev30histManagement") #,"rev18histManagement_ResAllom_2021_02_01")
outputdir <- "Output/"

@@

years             <- findset("past")
climateIPCCr      <- readRDS("Data/rev30histManagement/IPCC_reduced.rds")

@@

# CShare_ipcczone   <- dimSums(climateIPCCr * SOCDens[,,"crop"][,,"actualstate"]   * Landuse[,years,"crop"], dim=1)/
#                     dimSums(climateIPCCr * SOCDens[,,"natveg"][,,"naturalstate"] * Landuse[,years,"crop"], dim=1)

CShare_ipcczone   <- dimSums(climateIPCCr * toolConditionalReplace(SOCDens[,,"crop"][,,"actualstate"]/SOCDens[,,"natveg"][,,"naturalstate"],
                                                                  c("is.na()", "is.nan()", "is.infinite()",">3"), c(1,1,1,5)) * Landuse[,years,"crop"], dim=1) /
                     dimSums(climateIPCCr * Landuse[,years,"crop"], dim=1)

getNames(CShare_ipcczone) <- gsub("_", " ", getNames(CShare_ipcczone))
getCells(CShare_ipcczone) <- "GLO"

@@

climateIPCCr       <- readRDS("Data/rev30histManagement/IPCC_reduced.rds")
climateIPCC        <- readRDS("Data/rev30histManagement/IPCC.rds")

@@

MShare_ipcc <- LShare_ipcc <-  HShare_ipcc <- new.magpie("GLO",NULL, outer(c("IPCC2006","IPCC2019"),
                                                                          getNames(climateIPCCr), paste, sep="."))

@@

comparison <- mbind(CShare_ipcc, 
                    round(add_dimension(setYears(collapseNames(CShare_ipcczone[,"y1975","naturalstate"]),NULL),
                                  nm="This Study.hist.1975",add="model.input.year"),2),
                    round(add_dimension(setYears(collapseNames(CShare_ipcczone[,"y2010","naturalstate"]),NULL),
                                  nm="This Study.hist.2010",add="model.input.year"),2))

@@

save_kable(save, file="Output/Images/TableSCF_comparison_new.png")
table
```

### 2022-06-14 SOCbudgetPaper/introduction.Rmd

Soil Organic Carbon (SOC), the amount of organic carbon stored in the Earth’s soil, is the largest terrestrial organic carbon pool. It exceeds the carbon in the atmospheric and vegetation pools multiple times [@batjes_total_1996]. Even small changes in processes affecting SOC lead therefore to substantial shifts in the terrestrial carbon cycle and influence the amount of CO~2~ in the atmosphere [@friedlingstein_global_2019; @minasny_soil_2017]. The specific amount of carbon stored in soils globally is quantified with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile [@batjes_total_1996; @sanderman_soil_2017].

Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles to determine SOC variations over space and time. Recent studies have focused on the evaluation of total SOC stocks of the world as well as on the spatial disaggregation of soil properties such as SOC content [@batjes_harmonized_2016; @hengl_soilgrids250m_2017; @fao_global_2018]. However, these studies often do not include human interventions, like land cover change and agricultural management, in their analysis. Compared to climatic and geological drivers, they alter terrestrial carbon pools over much shorter time scales and are currently one of most dominant drivers of SOC changes on managed land.

The anthropogenic impact can be measured by the SOC debt (also referred to as SOC component of land-use change emissions), which is the amount of organic carbon soils have lost under cultivation compared to a potential natural vegetated state. Sanderman et al. [-@sanderman_soil_2017] identified the anthropogenic SOC debt for the first meter of the soil profile due to land cover change at around 116 GtC (37 GtC for the first 30 cm), compared to previous estimates of 60--130 GtC for the first meter [@lal_world_2001].

Global assessments of the carbon cycle via dynamic global vegetation models (DGVMs) and Earth System Models (ESMs) or bookkeeping models (BKMs) have analyzed SOC losses as part of a comprehensive evaluation of the global carbon budget and land-use change (LUC) emissions [@friedlingstein_global_2019]. While providing estimates of the magnitude of SOC losses due to land cover change, most models lack a detailed consideration of agricultural management. Earlier DGVM and ESM based assessments only considered changes in land cover, but ignored the removal of biomass at harvest [@strassmann_simulating_2008; @betts_climate_2015]. BKMs are designed to estimate LUC related emissons but ignore changes in SOC due to climate change, CO~2~ fertilization, and agricultural management [@friedlingstein_global_2019; @houghton_carbon_2012; @hansis_relevance_2015].

Managed agricultural systems were introduced in greater detail to DGVMs and ESMs to improve the assessment of the terrestrial carbon balance [e.g. @bondeau_modelling_2007; @lindeskog_implications_2013]. Pugh et al. [-@pugh_simulated_2015] explicitly consider agricultural management in the form of tillage, irrigation and biomass extraction at harvest, but worked with stylized scenarios rather than with historic management data. They also showed the importance of accounting for the land-use history, as many carbon emissions from agricultural soils are caused by historic LUC and the slow decline of SOC under cropland before it reaches a new equilibrium.

In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g. growing seasons [@portmann_mirca2000global_2010] multiple cropping systems [@waha_multiple_2020], irrigation systems [@jagermeyr_water_2015]) or as stylized scenarios [e.g. @pugh_simulated_2015, @lutz_simulating_2019].

More data sets on spatially explicit agricultural management time series with global coverage become available (e.g. on tillage systems, see [@porwollik_generating_2018], [@prestele_spatially_2018]) and model approaches are increasingly developed to project the dynamics of management systems into the future (e.g. [@iizumi_modeling_2019], [@minoli_modelling_2019]), but have --- to our knowledge --- not found their way into comprehensive assessments of the terrestrial carbon cycle in DGVMs and BKMs.

Field-scale models [@del_grosso_simulated_2001, @coleman_simulating_1997, @smith_estimating_2010, @taghizadeh-toosi_c-tool_2014] are able to better account for historic agricultural management if detailed information on crop yield levels, fertilizer inputs and various other on-farm measures is available for the studied sites. However, due to the lack of comprehensive global management data as input to these models, scaling up to the global domain remains a complex challenge.

The objective of our study is to provide the first global, spatially explicit SOC and SOC debt map that considers spatially explicit and time-variant historic agricultural management. To achieve this objective we create and provide a comprehensive data set of the global gridded management time series data, including crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices. We simulate SOC stocks, dynamics and the SOC debt for 1975--2010. Using a scenario analysis, we decompose the contribution of different management activities.

>>>>

Soil organic carbon (SOC), the amount of organic carbon stored in the Earth’s soil, constitutes the largest terrestrial organic carbon pool. It exceeds the carbon in the atmospheric and vegetation pools multiple times [@batjes_total_1996]. Even small changes in processes affecting SOC lead therefore to substantial shifts in the terrestrial carbon cycle and influence the amount of CO~2~ in the atmosphere [@frielingstein_glob_carbon_2020; @minasny_soil_2017]. The specific amount of carbon stored in soils globally is quantified with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile [@batjes_total_1996; @sanderman_soil_2017].

Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles to determine SOC variations over space and time. Recent studies have focused on the evaluation of total SOC stocks of the world as well as on the spatial disaggregation of soil properties such as SOC content [@batjes_harmonized_2016; @hengl_soilgrids250m_2017; @fao_global_2018]. However, these studies often do not include human interventions, like land cover change and agricultural management, in their analysis. Compared to climatic and geological driving forces, human interventions alter terrestrial carbon pools over much shorter time scales and are currently one of the most dominant drivers of SOC changes on managed land [@hansis_relevance_2015; @bastos_comparison_2021].

The anthropogenic impact can be measured by the SOC debt (also referred to as SOC component of land-use change emissions, see Pongratz et al. [-@pongratz_terminolgy_2014]), which is the amount of organic carbon soils have lost under cultivation compared to a hypothetical potential natural vegetation state. Sanderman et al. [-@sanderman_soil_2017] identified the anthropogenic SOC debt for the first meter of the soil profile due to land cover change at around 116 GtC (37 GtC for the first 30 cm), compared to previous estimates of 60--130 GtC for the first meter [@lal_world_2001].

Global assessments of the carbon cycle via dynamic global vegetation models (DGVMs), Earth System Models (ESMs) or bookkeeping models (BKMs) have analyzed SOC losses as part of a comprehensive evaluation of the global carbon budget and land-use change (LUC) emissions [@frielingstein_glob_carbon_2020]. While providing estimates of the magnitude of SOC losses due to land-cover change, most models lack a detailed representation of agricultural management. Earlier DGVM- and ESM-based assessments only considered changes in land cover, but ignored the removal of biomass at harvest [@strassmann_simulating_2008; @betts_climate_2015]. BKMs are designed to estimate LUC related emissions but often ignore changes in SOC due to climate change, CO~2~ fertilization and N deposition. Whereas BKM have largely improved in estimating additional emissions from wood harvest and shifting cultivation, state of the art models do not consider impacts of varying agricultural management [@frielingstein_glob_carbon_2020; @houghton_carbon_2012; @hansis_relevance_2015; @bastos_comparison_2021].

Managed agricultural systems were introduced in greater detail to DGVMs and ESMs to improve the assessment of the terrestrial carbon balance [e.g. @bondeau_modelling_2007; @lindeskog_implications_2013]. Pugh et al. [-@pugh_simulated_2015] explicitly consider agricultural management in the form of tillage, irrigation and biomass extraction at harvest, but worked with stylized scenarios rather than with historical management data. They also showed the importance of accounting for the land-use history, as many carbon emissions from agricultural soils are caused by historical LUC and the slow decline of SOC under cropland before a new equilibrium is reached.

In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g. growing seasons [@portmann_mirca2000global_2010], multiple cropping systems [@waha_multiple_2020], irrigation systems [@jagermeyr_water_2015]) or as stylized scenarios [e.g. @pugh_simulated_2015; @lutz_simulating_2019]. Herzfeld et al. [-@herzfeld_soc_2021] account for historical changes in fertilizer and manure inputs, residue removal rates and tillage systems and report SOC losses from cropland expansion over the period from 1700--2018 of 215 GtC. Within their stylized future management scenarios they find that none of the management aspects considered (residue management, no-tillage) can create a net carbon sink on current cropland areas under future climate change.

More data sets on spatially explicit agricultural management time series with global coverage are becoming available (e.g. on tillage systems, see [@porwollik_generating_2019; @prestele_spatially_2018]) and modeling approaches are increasingly being developed to project the dynamics of management systems into the future (e.g. [@iizumi_modeling_2019; @minoli_modelling_2019]), but have --- to our knowledge --- not yet found their way into comprehensive assessments of the terrestrial carbon cycle in DGVMs and BKMs.

Field-scale models [@del_grosso_simulated_2001; @coleman_simulating_1997; @smith_estimating_2010; @taghizadeh-toosi_c-tool_2014] are able to better account for historical agricultural management if detailed information on crop yield levels, fertilizer inputs and various other on-farm measures is available for the studied sites. However, due to the lack of comprehensive global management data as input to these models, scaling up to the global domain remains a complex challenge [@morais_detailed_2019].

Managed soils have been increasingly studied not only for their carbon emitting behavior, but also because of their capacity to re-store carbon (soil carbon sequestration (SCS) techniques). However, assessing SCS dynamically considering the interdependency with environmental, social and economic sustainability targets has been difficult so far, as integrated assessment models (IAMs) [@popp_land-use_2016; @rogelj_scenarios_2018; @forster_sr1p5_2018] have not integrated soil management into their mitigation pathways. More detailed process-based models are typically computationally too demanding to be integrated into optimization-based IAMs. Better accounting for soil carbon management in IAMs thus requires a light-weight model suitable for iterative modeling with detailed options to represent agricultural soil management.

The objectives of our study are (1) to develop a reduced-complexity SOC model able to account for SCS in IAM frameworks; (2) to create a comprehensive data set of the global gridded management time series, including crop production levels, residue input rates, manure amendments, and the adoption of irrigation and tillage practices; and (3) to provide global as well as spatially explicit SOC and SOC debt estimates that consider spatially explicit and time-variant agricultural management. 
We decompose the contribution of different management activities through a scenario analysis, identifying the most impacting management decisions for SOC development. Moreover, we compare our model performance against other SOC stock and SOC emission estimates, to evaluate the suitability of this reduced-complexity approach for integration into IAM modeling.
